---
title: "serac"
author: Rosalie Bruel, Pierre Sabatier
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Put the title of your vignette here}
  %\VignetteEngine{knitr::rmarkdown}
  \usepackage[utf8]{inputenc}
---

```{r, include=FALSE}
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L)
library(serac)
```

# First time user - setting up the folders

Which folder are your currently working in?

```{r eval=FALSE}
getwd()
```

If you want to work in another folder, change it with setwd(""), e.g.:

```{r eval=FALSE}
setwd("/myProject/Age depth model")
```


Line below will create the 'Cores' folder if it doesn't exist already
```{r eval=FALSE}
dir.create(file.path(getwd(), 'Cores'), showWarnings = FALSE)
```

Within the 'Cores' folder, user needs to create one folder per core.
```{r eval=FALSE}
dir.create(file.path(paste(getwd(),'/Cores',sep=""), 'serac_example_ALO09P12'), showWarnings = FALSE)

```

We are writing in this subfolder an example dataset.

Format your own data following this template, or get help with formatting using the serac_input_formatting() function

```{r eval=FALSE}
?serac_example_ALO09P12
write.table(x = serac_example_ALO09P12, file = paste(getwd(),'/Cores/serac_example_ALO09P12/serac_example_ALO09P12.txt',sep=""),col.names = T, row.names = F,sep="\t")

# Including proxy data for this core too
write.table(x = serac_example_ALO09P12_proxy, file = paste(getwd(),'/Cores/serac_example_ALO09P12/serac_example_ALO09P12_proxy.txt',sep=""),col.names = T, row.names = F,sep="\t")

```





# Now run serac
At the minimum, your function will be need the name of the core and the coring year, e.g.:

```{r eval=FALSE}
serac(name="serac_example_ALO09P12",coring_yr=2009)
```


Several sedimentation hypotheses were tested, and this is what the author chose as the best model:
```{r eval=FALSE}
serac(name="serac_example_ALO09P12",coring_yr=2009,model=c("CFCS"),plotphoto=FALSE,minphoto=c(0),maxphoto=c(210),plot_Pb=T,plot_Am=T,plot_Cs=T,Cher=c(30,40),Hemisphere=c("NH"),NWT=c(51,61),sedchange=c(75.5),plot_Pb_inst_deposit=T,inst_deposit=c(20,28,100,107,135,142,158,186),suppdescriptor=TRUE,descriptor_lab=c("Ca/Fe"),historic_d=c(20,28,100,107,135,142,158,186),historic_a=c(1994,1920,1886,1868),historic_n=c("sept1 994 flood","1920 flood","1886 flood","1868 flood ?"), min_yr=c(1750),dmax=c(180), plotpdf=TRUE,preview=F)

```

If you want to see the preview, change preview=T in the code.
Before that, make sure to extend your window in RStudio (large plot!)




# Outputs
Files with age-depth model and metadata are automatically saved in your working folder. If you assign your function to an object, serac will return a list with the parameters.

```{r eval=FALSE}
mymodel1 <- serac(name="serac_example_ALO09P12",coring_yr=2009)

# See the objects created
names(mymodel1)

# Visualize the input data
mymodel1$data

# Save the plot with custom parameters, e.g., Courier font and grey scale without editing default colors.
pdf(paste0(getwd(),"/Cores/serac_example_ALO09P12/serac_example_ALO09P12_custom.pdf"),width = 10, height = 5, family="Courier", colormodel = "grey")
mymodel1$plot
dev.off()

```

