---
title: "serac"
author: "Rosalie Bruel and Pierre Sabatier"
date: "Last update: `r Sys.Date()`"
output: 
  rmarkdown::html_vignette:
    toc: yes
    toc_depth: 3
    number_sections: true
vignette: >
  %\VignetteIndexEntry{serac}
  \usepackage[utf8]{inputenc}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include=FALSE}
stopifnot(require(knitr))
knitr::opts_chunk$set(collapse = T, comment = "#>")
options(tibble.print_min = 4L, tibble.print_max = 4L,width = 90)
opts_chunk$set(
  comment = NA,
  message = FALSE,
  warning = FALSE,
  dev = "png",
  dpi = 150,
  fig.asp = 0.8,
  fig.width = 5,
  out.width = "100%",
  fig.align = "center"
)

library(devtools)
devtools::install_github("rosalieb/serac")
library(serac)

```

The present vignette helps you compute your first age-depth model with _serac_, a R package for ShortlivEd RAdionuclide Chronology of recent sediment cores.

# First time user

## Setting up the folders

Which folder are your currently working in?

```{r getwd, eval=FALSE}
getwd()
```

If you want to work in another folder, change it with setwd(""), e.g.:

```{r setwd, eval=FALSE}
setwd("~/myProject/Age depth model")
```


Line below will create the 'Cores' folder if it doesn't exist already
```{r create core folder,  eval=TRUE}
dir.create(file.path(getwd(), 'Cores'), showWarnings = FALSE)
```

Within the 'Cores' folder, user needs to create one folder per core.
```{r create example data folder, eval=TRUE}
dir.create(file.path(paste(getwd(),'/Cores',sep=""), 'serac_example_ALO09P12'), showWarnings = FALSE)

```

We are writing in this subfolder an example dataset (Source: <a href:"https://www.sciencedirect.com/science/article/pii/S0033589412000294">Wilhelm et al., 2012</a>.

Format your own data following this template, or get help with formatting using the serac_input_formatting() function

```{r write example data, eval=TRUE}
?serac_example_ALO09P12
write.table(x = serac_example_ALO09P12, file = paste(getwd(),'/Cores/serac_example_ALO09P12/serac_example_ALO09P12.txt',sep=""),col.names = T, row.names = F,sep="\t")

# Including proxy data for this core too
write.table(x = serac_example_ALO09P12_proxy, file = paste(getwd(),'/Cores/serac_example_ALO09P12/serac_example_ALO09P12_proxy.txt',sep=""),col.names = T, row.names = F,sep="\t")

```


## User metadata (optional)
One of our goal with _serac_ is to improve reproducibility and data management. <br>
A first and easy step is to associate the name of the scientist (you!) to each model built. To enter personal data (recommended: name, affiliation, ORCID number and email), tape in the following line in the console, and follow instructions.

```{r user_infos,eval=FALSE, message=FALSE, warning=FALSE}
user_infos()
```



# Run serac
At the minimum, your function will be need the name of the core and the coring year. In the example below, I included the **save_code** argument because Rmarkdown doesn't save code history the way the code can use it. Unless you're including a plot in a Rmarkdown, I recommend you do not specify this argument, or turn it to its default, i.e., TRUE. That way, you can go back and see your code.

```{r model 1, eval=TRUE}
serac(name="serac_example_ALO09P12",coring_yr=2009,
      save_code=FALSE) #do not include this in the code if you're working outside a Rmarkdown document
```

Several sedimentation hypotheses were tested, and this is what the authors (<a href="https://www.sciencedirect.com/science/article/pii/S0033589412000294">Wilhelm _et al_, 2012</a>) chose as the best model:  
```{r model 2, eval=TRUE}
# Code:
model_ALO09P12 <-
  serac(name="serac_example_ALO09P12", coring_yr=2009, model=c("CFCS"),
      plotphoto=FALSE, minphoto=c(0), maxphoto=c(210),
      plot_Pb=T, plot_Am=T, plot_Cs=T, Cher=c(30,40), Hemisphere=c("NH"), NWT=c(51,61),
      sedchange=c(75.5),
      plot_Pb_inst_deposit=T, inst_deposit=c(20,28,100,107,135,142,158,186),
      suppdescriptor=TRUE, descriptor_lab=c("Ca/Fe"),
      historic_d=c(20,28,100,107,135,142,158,186),
      historic_a=c(1994,1920,1886,1868),
      historic_n=c("sept 1994 flood","1920 flood","1886 flood","1868 flood ?"), 
      min_yr=c(1750),
      dmax=c(180), 
      plotpdf=T, preview=T,
      save_code=FALSE)

# Output:
```

If you want to see the preview, change preview=T in the code.
Before that, make sure to extend your window in RStudio (large plot!)

```{r view final plot, echo=FALSE, message=FALSE, warning=FALSE}
model_ALO09P12$plot
```



# Explore outputs
Files with age-depth model and metadata are automatically saved in your working folder. If you assign your function to an object, serac will return a list with the parameters.

```{r save output, eval=TRUE}
# Code:
mymodel1 <- serac(name="serac_example_ALO09P12", coring_yr=2009, preview = T,save_code=FALSE)

# Output:
```

As always when computing a new code, _serac_ will check whether you tried the code combination previously. Here, we used the same code than in the previous section, so the code display this message.

Several objects are accessible from **mymodel1**:
`r paste("</br>    - ",names(mymodel1))`



```{r explore outputs, eval=TRUE, message=FALSE, warning=FALSE}
mymodel1$plot 

# Visualize the input data
knitr::kable(mymodel1$data, format = "markdown")

# Visualize the inventories (if density was in the input data file)
mymodel1$Inventories

# Save the plot with custom parameters, e.g., Courier font and grey scale without editing default colors.
pdf(paste0(getwd(),"/Cores/serac_example_ALO09P12/serac_example_ALO09P12_custom.pdf"),width = 10, height = 5, family="Courier", colormodel = "grey")
mymodel1$plot
dev.off()

```

# Other functions

## help_serac()
Another way to read some help.

## serac_input_formatting() 
If you unsure about the input format of your core (if you have depth in mm, or column names that don't follow our example), you can use the function *serac_input_formatting(name = "MyCore")* (MyCore being the name of your core and folder in which the data are located) and follow the steps.

## core_metadata()
Enter here a more extensive list of parameters associated to your core. Current list of parameters include:<br>
* ISGN <br>
* sample date (YYYY-MM-DD) <br>
* latitude (in WGS84)
* longtiude (in WGS84)
* coring method (e.g., gravity corer, piston corer, percussion, etc.) <br>
* subsampling method at the laboratory (e.g., calibrated volumetric sampler, extrusion, etc.) <br>
* instrument type (e.g., alpha spectrometry, well-type germanium detector, P-type germanium detector) <br>
* start of the measurements (YYYY-MM-DD) <br>
* end of the measurements (YYYY-MM-DD) <br>
* any additional comment (e.g., 210Pb background reached, first centimetres lost during coring, etc.) <br


## serac_map()
If you entered GPS coordinates for your core in the previous step, tape in *serac_map()* into your console and view the location of all the systems you entered GPS coordinates for. If you enter a vector, e.g., *serac_map(c("Lake1", "Lake2", "Lagoon1"))*, then the map will be generated solely for these systems.


```{r remove folder at the end, include=FALSE}
# remove folder that was created for compilation of this vignette.
unlink(paste0(getwd(),"/Cores"), recursive = TRUE)
```

```{r build_vigentte, eval=FALSE, include=FALSE}
# copy and run that line in the console to build vignette
#devtools::build_vignettes()
```

